```{r}
# importing necessary libraries
library(biomaRt)
library(org.Ce.eg.db)
library(clusterProfiler)
library(tidyverse)
library(GOSemSim)
library(enrichplot) 

```

```{r}
# Load the selected ribo and rna gene names
ribo_gene_name <- read.csv("selected_ribo.csv", row.names = 1)
rna_gene_name <- read.csv("selected_rna.csv", row.names = 1)

ribo_gene_name <- rownames(ribo_gene_name)
rna_gene_name <- rownames(rna_gene_name)

# Convert Ensembl IDs to gene symbols
ensembl <- useEnsembl(biomart = "ensembl", dataset = "celegans_gene_ensembl", version = 113)

ribo_wormbase_id <- getBM(attributes = c("ensembl_gene_id", "external_gene_name"),
                        filters = "external_gene_name",
                        values = ribo_gene_name,
                        mart = ensembl)

rna_wormbase_id <- getBM(attributes = c("ensembl_gene_id", "external_gene_name"),
                        filters = "external_gene_name",
                        values = rna_gene_name,
                        mart = ensembl)


head(ribo_wormbase_id)
dim(ribo_wormbase_id)
head(rna_wormbase_id)
dim(rna_wormbase_id)

# save the ids

write.csv(ribo_wormbase_id, file = "ribo_wormbase_id.csv", row.names = FALSE)
write.csv(rna_wormbase_id, file = "rna_wormbase_id.csv", row.names = FALSE)


```

```{r}
# Load the wormbase ids
ribo_wormbase_id <- ribo_wormbase_id$ensembl_gene_id
rna_wormbase_id <- rna_wormbase_id$ensembl_gene_id

# Convert WORMBASE IDs to ENTREZ IDs using org.Ce.eg.db
entrez_id_ribo <- mapIds(org.Ce.eg.db, keys = ribo_wormbase_id, column = "ENTREZID", keytype = "WORMBASE", multiVals = "list")

entrez_id_rna <- mapIds(org.Ce.eg.db, keys = rna_wormbase_id, column = "ENTREZID", keytype = "WORMBASE", multiVals = "list")

entrez_id_ribo_df <- data.frame(WB_ID = rep(names(entrez_id_ribo), sapply(entrez_id_ribo, length)),
                            ENTREZ_ID = unlist(entrez_id_ribo))

entrez_id_rna_df <- data.frame(WB_ID = rep(names(entrez_id_rna), sapply(entrez_id_rna, length)),
                            ENTREZ_ID = unlist(entrez_id_rna))

head(entrez_id_ribo_df)
dim(entrez_id_ribo_df)
head(entrez_id_rna_df)

dim(entrez_id_rna_df)

# save the ids
write.csv(entrez_id_ribo_df, file = "entrez_id_ribo_df.csv", row.names = FALSE)
write.csv(entrez_id_rna_df, file = "entrez_id_rna_df.csv", row.names = FALSE)

```

```{r}
# Load the wormbase ids
ribo_wormbase_id <- read.csv("ribo_wormbase_id.csv")
rna_wormbase_id <- read.csv("rna_wormbase_id.csv")
ribo_entrez_id <- read.csv("entrez_id_ribo_df.csv")
rna_entrez_id <- read.csv("entrez_id_rna_df.csv")

# renaming the columns
colnames(ribo_wormbase_id) <- c("WB_ID", "external_gene_name")
colnames(rna_wormbase_id) <- c("WB_ID", "external_gene_name")

head(ribo_wormbase_id)
head(ribo_entrez_id)
dim(ribo_wormbase_id)
dim(ribo_entrez_id)

#merge the two dataframes
new_df_ribo <- merge(ribo_wormbase_id, ribo_entrez_id, by = "WB_ID", all.x = TRUE)
new_df_rna <- merge(rna_wormbase_id, rna_entrez_id, by = "WB_ID", all.x = TRUE)
head(new_df_ribo)
head(new_df_rna)
dim(new_df_ribo)
dim(new_df_rna)
# save the ids
write.csv(new_df_ribo, file = "all_ribo_id.csv", row.names = FALSE)
write.csv(new_df_rna, file = "all_rna_id.csv", row.names = FALSE)
```


```{r}

setdiff(ribo_gene_name,ribo_wormbase_id$external_gene_name) # check if there are any missing genes in the ribo data

```

```{r}
# Load the Entrez IDs for ribo and rna
entrez_id_ribo_df <- read.csv("entrez_id_ribo_df.csv")
entrez_id_rna_df <- read.csv("entrez_id_rna_df.csv")


```{r}
# Perform GO enrichment analysis for ribo data (Biological Process, Molecular Function, Cellular Component)
go_enrich_ribo_bp <- enrichGO(gene = entrez_id_ribo_df$ENTREZ_ID,
                     OrgDb = org.Ce.eg.db,
                     keyType = "ENTREZID",
                     ont = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff = 0.05)
go_enrich_ribo_mf <- enrichGO(gene = entrez_id_ribo_df$ENTREZ_ID,
                     OrgDb = org.Ce.eg.db,
                     keyType = "ENTREZID",
                     ont = "MF",
                     pAdjustMethod = "BH",
                     pvalueCutoff = 0.05)
go_enrich_ribo_cc <- enrichGO(gene = entrez_id_ribo_df$ENTREZ_ID,
                     OrgDb = org.Ce.eg.db,
                     keyType = "ENTREZID",
                     ont = "CC",
                     pAdjustMethod = "BH",
                     pvalueCutoff = 0.05)



go_enrich_ribo_bp_result <- go_enrich_ribo_bp@result
go_enrich_ribo_mf_result <- go_enrich_ribo_mf@result
go_enrich_ribo_cc_result <- go_enrich_ribo_cc@result                  

```

```{r}
# Perform GO enrichment analysis for rna data (Biological Process, Molecular Function, Cellular Component)
go_enrich_rna_bp <- enrichGO(gene = entrez_id_rna_df$ENTREZ_ID,
                     OrgDb = org.Ce.eg.db,
                     keyType = "ENTREZID",
                     ont = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff = 0.05)
go_enrich_rna_mf <- enrichGO(gene = entrez_id_rna_df$ENTREZ_ID,
                        OrgDb = org.Ce.eg.db,
                        keyType = "ENTREZID",
                        ont = "MF",
                        pAdjustMethod = "BH",
                        pvalueCutoff = 0.05)
go_enrich_rna_cc <- enrichGO(gene = entrez_id_rna_df$ENTREZ_ID,
                        OrgDb = org.Ce.eg.db,
                        keyType = "ENTREZID",
                        ont = "CC",
                        pAdjustMethod = "BH",
                        pvalueCutoff = 0.05)


go_enrich_rna_bp_result <- go_enrich_rna_bp@result
go_enrich_rna_mf_result <- go_enrich_rna_mf@result
go_enrich_rna_cc_result <- go_enrich_rna_cc@result
```

```{r}
dim(go_enrich_ribo_bp_result)
dim(go_enrich_ribo_mf_result)
dim(go_enrich_ribo_cc_result)
dim(go_enrich_rna_bp_result)
dim(go_enrich_rna_mf_result)
dim(go_enrich_rna_cc_result)


```

```{r}
# Remove NA values from the GO enrichment results
go_enrich_ribo_bp_result_na <- na.omit(go_enrich_ribo_bp_result)
go_enrich_ribo_mf_result_na <- na.omit(go_enrich_ribo_mf_result)
go_enrich_ribo_cc_result_na <- na.omit(go_enrich_ribo_cc_result)
go_enrich_rna_bp_result_na <- na.omit(go_enrich_rna_bp_result)
go_enrich_rna_mf_result_na <- na.omit(go_enrich_rna_mf_result)
go_enrich_rna_cc_result_na <- na.omit(go_enrich_rna_cc_result)

dim(go_enrich_ribo_bp_result_na)
dim(go_enrich_ribo_mf_result_na)
dim(go_enrich_ribo_cc_result_na)
dim(go_enrich_rna_bp_result_na)
dim(go_enrich_rna_mf_result_na)
dim(go_enrich_rna_cc_result_na)

```
```{r}
# extracting the GO ID's for the terms(Ribo)
go_terms_ribo_bp <- go_enrich_ribo_bp_result_na$ID
go_terms_ribo_mf <- go_enrich_ribo_mf_result_na$ID
go_terms_ribo_cc <- go_enrich_ribo_cc_result_na$ID

# using the ID's to extract the terms data
go_data_ribo_bp <- godata("org.Ce.eg.db", ont = "BP") 
go_data_ribo_mf <- godata("org.Ce.eg.db", ont = "MF")
go_data_ribo_cc <- godata("org.Ce.eg.db", ont = "CC")

#checking for the similarity within the terms 
sim_matrix_ribo_bp <- mgoSim(go_terms_ribo_bp, go_terms_ribo_bp ,semData = go_data_ribo_bp, measure = "Rel", combine = NULL)
sim_matrix_ribo_mf <- mgoSim(go_terms_ribo_mf, go_terms_ribo_mf ,semData = go_data_ribo_mf, measure = "Rel", combine = NULL)
sim_matrix_ribo_cc <- mgoSim(go_terms_ribo_cc, go_terms_ribo_cc ,semData = go_data_ribo_cc, measure = "Rel", combine = NULL)

sim_matrix_ribo_bp_df <- as.data.frame(as.table(sim_matrix_ribo_bp))
sim_matrix_ribo_mf_df <- as.data.frame(as.table(sim_matrix_ribo_mf))
sim_matrix_ribo_cc_df <- as.data.frame(as.table(sim_matrix_ribo_cc))
```

```{r}
# extracting the GO ID's for the terms(Rna)
go_terms_rna_bp <- go_enrich_rna_bp_result_na$ID
go_terms_rna_mf <- go_enrich_rna_mf_result_na$ID
go_terms_rna_cc <- go_enrich_rna_cc_result_na$ID

# using the ID's to extract the terms data
go_data_rna_bp <- godata("org.Ce.eg.db", ont = "BP")
go_data_rna_mf <- godata("org.Ce.eg.db", ont = "MF")
go_data_rna_cc <- godata("org.Ce.eg.db", ont = "CC")
#checking for the similarity within the terms 
sim_matrix_rna_bp <- mgoSim(go_terms_rna_bp, go_terms_rna_bp ,semData = go_data_rna_bp, measure = "Rel", combine = NULL)
sim_matrix_rna_mf <- mgoSim(go_terms_rna_mf, go_terms_rna_mf ,semData = go_data_rna_mf, measure = "Rel", combine = NULL)
sim_matrix_rna_cc <- mgoSim(go_terms_rna_cc, go_terms_rna_cc ,semData = go_data_rna_cc, measure = "Rel", combine = NULL)

sim_matrix_rna_bp_df <- as.data.frame(as.table(sim_matrix_rna_bp))
sim_matrix_rna_mf_df <- as.data.frame(as.table(sim_matrix_rna_mf))
sim_matrix_rna_cc_df <- as.data.frame(as.table(sim_matrix_rna_cc))
```
```{r}
 # selecting the highly similar terms
highly_similar_ribo_bp <-  subset(sim_matrix_ribo_bp_df, Freq > 0.95 & Var1 != Var2) 
highly_similar_ribo_mf <-  subset(sim_matrix_ribo_mf_df, Freq > 0.95 & Var1 != Var2)
highly_similar_ribo_cc <-  subset(sim_matrix_ribo_cc_df, Freq > 0.95 & Var1 != Var2)

highly_similar_rna_bp <-  subset(sim_matrix_rna_bp_df, Freq > 0.95 & Var1 != Var2)
highly_similar_rna_mf <-  subset(sim_matrix_rna_mf_df, Freq > 0.95 & Var1 != Var2)
highly_similar_rna_cc <-  subset(sim_matrix_rna_cc_df, Freq > 0.95 & Var1 != Var2)
```
```{r}
# Removing redundant terms
keep_ribo_bp <- go_enrich_ribo_bp_result_na %>% arrange(p.adjust) %>%
  filter(!(ID %in% highly_similar_ribo_bp$Var2))  

keep_ribo_mf <- go_enrich_ribo_mf_result_na %>% arrange(p.adjust) %>%
    filter(!(ID %in% highly_similar_ribo_mf$Var2))
keep_ribo_cc <- go_enrich_ribo_cc_result_na %>% arrange(p.adjust) %>%
    filter(!(ID %in% highly_similar_ribo_cc$Var2))

keep_rna_bp <- go_enrich_rna_bp_result_na %>% arrange(p.adjust) %>%
  filter(!(ID %in% highly_similar_rna_bp$Var2))
keep_rna_mf <- go_enrich_rna_mf_result_na %>% arrange(p.adjust) %>%
    filter(!(ID %in% highly_similar_rna_mf$Var2))

keep_rna_cc <- go_enrich_rna_cc_result_na %>% arrange(p.adjust) %>%
    filter(!(ID %in% highly_similar_rna_cc$Var2))
```





```{r}
# Create a new GO enrichment object with filtered terms

go_enrich_ribo_bp_filtered <- go_enrich_ribo_bp
go_enrich_ribo_mf_filtered <- go_enrich_ribo_mf
go_enrich_ribo_cc_filtered <- go_enrich_ribo_cc
go_enrich_rna_bp_filtered <- go_enrich_rna_bp
go_enrich_rna_mf_filtered <- go_enrich_rna_mf   
go_enrich_rna_cc_filtered <- go_enrich_rna_cc

go_enrich_ribo_bp_filtered@result <- keep_ribo_bp
go_enrich_ribo_mf_filtered@result <- keep_ribo_mf
go_enrich_ribo_cc_filtered@result <- keep_ribo_cc
go_enrich_rna_bp_filtered@result <- keep_rna_bp
go_enrich_rna_mf_filtered@result <- keep_rna_mf
go_enrich_rna_cc_filtered@result <- keep_rna_cc

```


```{r}
# Save the filtered GO enrichment results to CSV files
write.csv(go_enrich_ribo_bp_filtered@result, file = "go_enrich_ribo_bp_filtered.csv", row.names = FALSE)
write.csv(go_enrich_ribo_mf_filtered@result, file = "go_enrich_ribo_mf_filtered.csv", row.names = FALSE)
write.csv(go_enrich_ribo_cc_filtered@result, file = "go_enrich_ribo_cc_filtered.csv", row.names = FALSE)
write.csv(go_enrich_rna_bp_filtered@result, file = "go_enrich_rna_bp_filtered.csv", row.names = FALSE)
write.csv(go_enrich_rna_mf_filtered@result, file = "go_enrich_rna_mf_filtered.csv", row.names = FALSE)
write.csv(go_enrich_rna_cc_filtered@result, file = "go_enrich_rna_cc_filtered.csv", row.names = FALSE)
```

```{r}

# # count is na
# go_enrich_ribo_bp_filtered@result$Count[is.na(go_enrich_ribo_bp_filtered@result$Count)] <- 0

head(go_enrich_ribo_bp_filtered@result)

``` 

```{r}
# Plot the dotplot for the filtered GO enrichment results
png("go_enrich_ribo_bp_filtered.png", width = 800, height = 600)  
dotplot(go_enrich_ribo_bp_filtered, showCategory = 10) + ggtitle("Ribo-seq GO Enrichment BP")
dev.off()
png("go_enrich_ribo_mf_filtered.png", width = 800, height = 600)
dotplot(go_enrich_ribo_mf_filtered, showCategory = 10) + ggtitle("Ribo-seq GO Enrichment MF")
dev.off()
png("go_enrich_ribo_cc_filtered.png", width = 800, height = 600)
dotplot(go_enrich_ribo_cc_filtered, showCategory = 10) + ggtitle("Ribo-seq GO Enrichment CC")
dev.off()
png("go_enrich_rna_bp_filtered.png", width = 800, height = 600)
dotplot(go_enrich_rna_bp_filtered, showCategory = 10) + ggtitle("RNA-seq GO Enrichment BP")
dev.off()
png("go_enrich_rna_mf_filtered.png", width = 800, height = 600)
dotplot(go_enrich_rna_mf_filtered, showCategory = 10) + ggtitle("RNA-seq GO Enrichment MF")
dev.off()
png("go_enrich_rna_cc_filtered.png", width = 800, height = 600)
dotplot(go_enrich_rna_cc_filtered, showCategory = 10) + ggtitle("RNA-seq GO Enrichment CC")
dev.off()

```

```{r}
go_enrich_ribo_bp_sim <- pairwise_termsim(go_enrich_ribo_bp_filtered)
go_enrich_ribo_mf_sim <- pairwise_termsim(go_enrich_ribo_mf_filtered)
go_enrich_ribo_cc_sim <- pairwise_termsim(go_enrich_ribo_cc_filtered)
go_enrich_rna_bp_sim <- pairwise_termsim(go_enrich_rna_bp_filtered)
go_enrich_rna_mf_sim <- pairwise_termsim(go_enrich_rna_mf_filtered)
go_enrich_rna_cc_sim <- pairwise_termsim(go_enrich_rna_cc_filtered)

```
```{r}
# Plot the enrichment map for the filtered GO enrichment results
png("go_enrich_ribo_bp_sim.png", width = 800, height = 600)
emapplot(go_enrich_ribo_bp_sim, showCategory = 10) + ggtitle("Ribo-seq GO Emapplot BP")
dev.off()
png("go_enrich_ribo_mf_sim.png", width = 800, height = 600)
emapplot(go_enrich_ribo_mf_sim, showCategory = 10) + ggtitle("Ribo-seq GO Emapplot MF")
dev.off()
png("go_enrich_ribo_cc_sim.png", width = 800, height = 600)
emapplot(go_enrich_ribo_cc_sim, showCategory = 10) + ggtitle("Ribo-seq GO Emapplot CC")
dev.off()
png("go_enrich_rna_bp_sim.png", width = 800, height = 600)
emapplot(go_enrich_rna_bp_sim, showCategory = 10) + ggtitle("RNA-seq GO Emapplot BP")
dev.off()
png("go_enrich_rna_mf_sim.png", width = 800, height = 600)
emapplot(go_enrich_rna_mf_sim, showCategory = 10) + ggtitle("RNA-seq GO Emapplot MF")
dev.off()
png("go_enrich_rna_cc_sim.png", width = 800, height = 600)
emapplot(go_enrich_rna_cc_sim, showCategory = 10) + ggtitle("RNA-seq GO Emapplot CC")
dev.off()
```